---
title: "Generics and Methods"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Function Overload

As far we know, the summary function have different output depending on what is the type of input. For instance, 

```{r cars}
#logical values
p1 <- c(NA, NA, NULL, NA)
summary(p1)

#numeric values
p2 <- 1:10
summary(p2)
```

Both outputs have different formats (heaeders, values) for each input.

Different kind of behavour according to different type of inputs is called

`Function Overloading`.

The `S3` exists to make this problem much more easier to deal with. This system split the function in two parts

1. Generic function
2. Method function (for each class)

An example of generic function is `print`. Let's see:

```{r}
print
```

This function is really simple, this is typical from `S3`generic, it just use `UseMethod` as string. 

## Condition for S3 Methods
There two condition to

1. The name of the `method` = `name of the generic`.`name of the class of the variable`. For example in R,

- `print.Date`
- `print.function`
- `print.factor`
- `summary.factor`
- `unique.array`

In the first case, `print` is the **name of the generic method** , `dot`, **the name of the class** `Date`.


2. The arguments of the **method**, must have all the arguments of the **generic**

For example the arguments of `print` is `x` and `...` or elipsis arguemtn. Hence, `print.Date` must have `x` and `...` too. 

```{r, eval=TRUE}
args(print)
```

```{r, eval=TRUE}
args(print.Date)
```

Furthermore, the elipsis arguments (`...`) allow to pass `arguments` between `methods`.

Sometimes, there are confusion between generic functions and other functions that people usually write. `S3` use `.` to specify `methods`. Therefore, a good practice is `camelCase` or `snake_case` functions names.

One way to check whether functions are `S3` methods is the `pryr` package.

```{r, warning=FALSE, message=FALSE}
library(pryr)
is_s3_generic("t")           # generic transpose function
is_s3_method("t.data.frame") # transpose method for data.frames
is_s3_method("t.test")       # a function for Student's t-tests
is_s3_method("is.na.data.frame") #is a method of the is.na generic that acts on data.frame objects.
```

The `S3` generics contain a call to `UseMethod`, and `S4` methods take the form `generic.class`.


## Creating a Generic Function

You can create your own `S3` functions. The first step is to write the generic. This is typically a single line function that calls `UseMethod()`, passing its name as a string.

The first argument to an `S3` generic is usually called x, though this isn't compulsory. It is also good practice to include a `...` ("ellipsis", or "dot-dot-dot") argument, in case arguments need to be passed from one method to another.

Overall, the structure of an `S3` generic looks like this.


```{r}
an_s3_generic <- function(x, maybe = "some", other = "arguments", ...) {
  UseMethod("an_s3_generic")
}
```

An example might be define an `S3` generic function to calculate the number of elements in an object `x`:
  
- Assign the function to `get_n_elements`.
- It should have arguments `x` and `....`
- The body of the function should call `UseMethod()`.


```{r}
# Create get_n_elements
get_n_elements<-function(x, ...){
  UseMethod("get_n_elements")
}
```

By itself, the generic function doesn't do anything. For that, you need to create methods, which are just regular functions with two conditions:

1. The name of the method must be of the form generic.class.
2. The method signature - that is, the arguments that are passed in to the method - must contain the signature of the generic.

The example explain step-by-step how to create a method using the generic `S3` function above:


-  View get_n_elements `generic`
```{r}
get_n_elements
```

- Create a data.frame `method` for get_n_elements
```{r}
get_n_elements.data.frame<-function(x,...){
  nrow(x)*ncol(x)
}
```

- Call the method on the sleep dataset
```{r}
n_elements_sleep <- get_n_elements(sleep) 
```

- View the result
```{r}
n_elements_sleep
```

## Default Method: Generalization for different type of data

If no suitable method is found for a generic, then an error is thrown. For example, at the moment, get_n_elements() only has a method available for data.frames. If you pass a matrix to get_n_elements() instead, you'll see an error.

```{r}
get_n_elements(matrix())
```

Rather than having to write dozens of methods for every kind of input, you can create a `method` that handles all types that don't have a specific method. This is called the **default** method; it always has the name generic.default. For example, `print.default()` will print any type of object that doesn't have its own `print()` method.

Do continue with the example:


- View pre-defined objects
```{r}
ls.str()
```

- Create a default method for get_n_elements
```{r}
get_n_elements.default<-function(x, ...){
  length(unlist(x))
}
```

- Call the method on the ability.cov dataset
```{r}
n_elements_ability.cov <- get_n_elements.default(ability.cov)
```


As we can se the `default` method catches all classes that don't have their own method.


